---
title: "RFM and CLV Shiny App funcs"
output:
  html_document:
    df_print: paged
---

For Lec04, Unsupervised Learning, this below is 1 of 3 planned apps. The other 2 being 

* a basic PCA app
* a recommendation systems app (extended to accept longform DF inputs)

## Basic Idea of RFM-CLV app

We read in a long-form DF that should contain at a minimum these 4 fields:

* a `customerID` column
* a `txn_ID` column such as Invoice number etc.
* a `txn_amount` column, and
* a `txn_date` column convertible to `as.Date(YY-MM-DD)`

Other variables can be around but won't matter as much. User will select those 3 variables via drop-downs in UI.

Then I use the `rfm` package to build a RFM data frame from long-form.

Now, to get CLV for each custID from the RFM wala DF, I need the following quantities:

* avg store `gross_margin` (UI default=10% or 0.1, & slider from 0.01 to 0.99)
* `discount_rate` (from UI, default=0.1 but slider goes from 0.1 to 0.50)
* `retention_rate` (calculated as below)
* a `churn_threshold` as percentile of days_since_last_visit (from UI, default=0.75)

To get retention_rate per cust_ID as a probability, I use logit from `stats::glm()`. The DV is a binary class (Y = churned or not) and the IVs are F and M from RFM.

Then, I plug the retention_rate into the following, [overly] simple infinite-horizon wala CLV formula and we're on:

* CLV  = Tot_rev_it * gross_margin * (ret_rate_it/(1 + discount_rate - ret_rate_it))

for custID indexed by i and t indexing last period (it uniquely IDs each row).

## Basic App UI design

Need the following Input fields on the main page:

* file input for txns file
* drop-down variables list for selecting CustID colm
* drop-down variables list for selecting txnID colm
* drop-down variables list for selecting txn_date colm
* drop-down variables list for selecting txn_amount colm

And output tabs could be:

* Intro and Overview
* Data summary tab with top 10 HTML tbl rows, DF size, missingness tbl
* RFM tab where we show `df0_rfm` top 10 rows and allow download
* CLV tab where we show `df1` top 10 rows and allow full DF download

On CLV output tab, we can collect more UI inputs, namely:

* churn_thresh_ui = 0.75  (default. slider from 0.5 to 0.99)
* discount_rate_ui = 0.1  (default, UI slider from 0.01 to 0.20)
* profit_margin_ui = 0.1  (default, else numeric input field)


## Setup chunk

```{r setup_load}

suppressPackageStartupMessages({
  if (!require('data.table')){install.packages("data.table")}; library("data.table")
  if (!require('tidyverse')){install.packages("tidyverse")}; library("tidyverse")
  if (!require('rfm')){install.packages("rfm")}; library("rfm")
  if (!require('stats')){install.packages("stats")}; library("stats")
})

# Load data
path0 = "D:/Dropbox/teaching related/MLBM/Co 2021/Lec04 Unsupervised Learning/"
df_data <- fread(paste0(path0, "ecommData_rfm_clv.csv"))
head(df_data)

```

Now, let me collect all UI inputs required from user at one place below.

```{r all_UI}

## UI inputs for RFM
custID_ui = as.factor(df_data$CustomerID)  # from UI, make `as.factor()`

txnID_ui = as.factor(df_data$InvoiceNo)  # from UI, make `as.factor()`

txn_amount_ui = as.numeric(df_data$totRev)  # from UI, make `as.numeric()`

date_format = "%m/%d/%Y %H:%M" # from UI. Tough but will ask folks do this. 

txn_date_ui = as.Date(df_data$InvoiceDate, date_format)

rfm_bins_num = 4 # from UI. Drop-down from 3 to 8




# UI inputs for CLV
churn_thresh_ui = 0.75  # default. slider from 0.5 to 0.99

discount_rate_ui = 0.1  # default, UI slider from 0.01 to 0.20

profit_margin_ui = 0.1 # default, else numeric input field

```

## Build RFM wala DF

Here, we cobble together user inputs from UI and build our main RFM wala DF.

I write a func that will do it in one go. behold.

```{r make_df_rfm}

## RFM builder func
get_rfm_df <- function(custID_ui=custID_ui, txnID_ui = txnID_ui,
                 txn_amount=txn_amount_ui, txn_date_ui = txn_date_ui,
		              rfm_bins_num=rfm_bins_num){
  
  df1 = data.frame(custID_ui=custID_ui, txnID_ui = txnID_ui,
                 txn_amount=txn_amount_ui, txn_date_ui = txn_date_ui)

  ## RFM calc
  df_RFM <- df1 %>% 
    group_by(custID_ui) %>% # all stuff below is by cust_ID
  
    summarise(
      recency = as.numeric(max(df1$txn_date_ui) - max(txn_date_ui)),
      freq = n_distinct(txnID_ui), 
      monetary= sum(txn_amount)/n_distinct(txnID_ui)) %>% # 12.5s 
  	# summary(df_RFM)
  	rfm_table_customer(.,
                           customer_id = custID_ui,
                           n_transactions = freq, 
                           recency_days = recency,
                           total_revenue = monetary, 
                           recency_bins=rfm_bins_num, 
                           frequency_bins = rfm_bins_num, 
                           monetary_bins = rfm_bins_num)
  
  df0_rfm = df_RFM$rfm
  return(df0_rfm)
	} # func ends

```

## Onto CLV analysis

Our Core CLV formula is: 

`CLV_it = (totRev_it * Profit_margin%_ui)*[r_it/(1+d_ui - r_it)]`

where `r_it` = retention rate = estimated from glm::logreg

`Pr(Y_it = 0 | X_it) = logit(X_it*beta)` for `X_it = {R_it, F_it, M_it}`

and `Y_it = 0` for `R_it >= some_thresh_ui`

Using UI based inputs from chunk above, I now proceed to build the binary DV (Y = churned or not) and thereafter, the analysis DF. 

Behold.

```{r clv_calc_func}

## CLV calc func
calc_clv <- function(data,
                     churn_thresh_ui = 0.75,
                     discount_rate_ui=0.10, 
                     profit_margin_ui=0.10){

  y = rep(0, nrow(data)); y[1:4]
    a0 = data$recency_days %>% quantile(., probs = churn_thresh_ui); #a0
    a1 = (data$recency_days >= a0); #a1[1:4]
    y[a1] = 1; #y[1:8]
  
  df0 = data.frame(y=y, 
                       recency0 = data$recency_days,
                       freq0 = data$transaction_count,
                       monetary0 = data$amount,
                       Rscore = data$recency_score,
                       Fscore=data$frequency_score,
                       Mscore = data$monetary_score)

  mylogit <- glm(y ~ freq0 + monetary0 + factor(Fscore) + factor(Mscore), 
                 data = df0, family = "binomial")  # 0.2s!

  summary(mylogit) # can display if needed

  a2 = mylogit$fitted.values
  ret_rate = 1 - a2; # ret_rate[1:8]
  
  A = ret_rate / (1 + discount_rate_ui - ret_rate); #A[1:8]
  clv = df0$monetary0 * profit_margin_ui * A; #clv[1:8]
  
  df1 = data.frame(df0, CLV=clv); print(head(df1))
  
  return(df1) # main output DF for CLV
  	} # func ends

```

## Test-drive the funcs and examine outputs

```{r test_drive}

system.time({ 
  
  df0_rfm = get_rfm_df(custID_ui=custID_ui, 
                       txnID_ui = txnID_ui,
                       txn_amount=txn_amount_ui, 
                       txn_date_ui = txn_date_ui,
                       rfm_bins_num=rfm_bins_num)
  
  print(head(df0_rfm))
  
	df1 = calc_clv(df0_rfm, 
	               churn_thresh_ui = churn_thresh_ui, 
	               discount_rate_ui = discount_rate_ui, 
	               profit_margin_ui = profit_margin_ui)
	
	print(head(df1))
	
	})
```


Dassit, we're done. Can make histograms etc for CLV thus obtained, if needed.

Slightly involved app in that requires multiple UI inputs from user but what to do only. 

Sudhir, June 2021.